---
title: "Final Project - SK4.1"
author: "Sarah Im" 
output: html_document
date: "2024-12-04"
bibliography: doc/myrefs.bib
---

# Do voters want Republicans or Democrats in Congress? 

## Background

Public opinion polls from different pollsters and organizations are used to demonstrate the performance and reputation of the political parties in our democratic society. 
The public approval ratings of the Democratic and Republican parties were measured using public opinion polls in this study. The voter behavior in the United States has evolved over the years and can be studied using public opinion polling. These statistics are often used for campaign strategies of different groups and as a tool in academics and journalism [@hillygus2011evolution]. This study utilizes data from FiveThirtyEight’s generic congressional ballots and focuses on the approval ratings of the democratic and republican parties in congress from November 21, 2020 to August 20, 2022 [@fivethirtyeight2022]. Analysis of these polls can reveal insights such as the changing voter preference and attitude in correspondence to key events such as election cycles, policy changes, and national crises. However, the credibility and biases over public opinion polling often changes when an individual views the polls. Some individuals view certain polls as “more reliable” if the polls align with their personal beliefs. In addition, correlation between party polarization and poll results can be found [@madson2020all]. The interpretation of the poll data can be essential for assessing the political dynamics in the United States and can reveal trends within voter alignment, policy approval, etc. 

This research focuses on the question: 

How have approval ratings between the Democratic and Republican parties in Congress changed over time? 

## Analysis 

In this study, the dataset was imported from FiveThirtyEight’s generic congressional ballots which contained voter ratings from November 2020 to November 2022. The data set was uploaded onto R studio, processed, and cleaned by standardizing the column names, selecting data points for the specific political party (dem or rep), and missing values were removed from the original data set. Furthermore, the dates from the original dataset were transformed and formatted to calculate the number of days since the origin date of the polls (11/21/2020) to allow for better visualization of data trends. 
For plotting the visualizations, the data was reshaped to plot the approval ratings of both parties separately. For the visualization for the democratic party, the data points are represented in blue and demonstrate the daily average approval ratings. For the visualization for the republican party, the data points are represented in red and demonstrate the daily average approval ratings. 

### 1. Load required libraries:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loading required libraries
library(tidyr)
library(janitor)
library(ggplot2)
library(dplyr)
```


### 2. Data Import and Data Cleaning of the Democratic Party 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.data <- read.csv("data/generic_polllist (1).csv")

ballot.Approve <- ballot.data %>%
  clean_names() %>%
  select(startdate, dem) %>%
  na.omit()

ballot.Approve$startdate <- as.Date(ballot.Approve$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve$days_since_origin <- as.numeric(ballot.Approve$startdate - origin_date)

ballot.Approve
```
### 3. Data Transformation: Conversion of the Data set 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve
ballot.wide <- pivot_wider(ballot.Approve, names_from = startdate, values_from = dem, values_fn=mean) 
ballot.wide

ballot.long <- pivot_longer(ballot.wide, cols = !days_since_origin, names_to = "startdate", values_to = "dem")
```


### 4. Visualization of the data using ggplot to represent the Democratic Approval Rates

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 1. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in approval ratings of the democratic party in congress over time since the origin of the polls on 11/21/20. The democratic party experienced a mean approval rating of around 43.9% with the range being approximately 11.3% (min. value = 38.3%, highest value=50.0%) from 11/21/20 to 11/05/2022."}

vis1 <- ggplot(ballot.long, aes(x=days_since_origin, y=dem)) + geom_point(color = "blue", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Democratic Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + ylim(35,50)
vis1
```

### 5. Data Cleaning of the Republican Party 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve2 <- ballot.data %>%
  clean_names() %>%
  select(startdate, rep) %>%
  na.omit()

ballot.Approve2$startdate <- as.Date(ballot.Approve2$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve2$days_since_origin <- as.numeric(ballot.Approve2$startdate - origin_date)
ballot.Approve2
```


### 6. Data Transformation: Conversion of the Data set  

```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve2
ballot.wide2 <- pivot_wider(ballot.Approve2, names_from = startdate, values_from = rep, values_fn=mean) 
ballot.long2 <- pivot_longer(ballot.wide2, cols = !days_since_origin, names_to = "startdate", values_to = "rep") 
```

### 7. Visualization of the data using ggplot to represent the Republican Approval Rates

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 2. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in average approval ratings of the republican party in congress over time since the origin of the polls on 11/21/20. The republican party experienced a mean approval rating of around 42.28% with the range being 14.7% (lowest rating = 35.0%, highest value = 49.7%) from 11/21/20 to 11/05/22."}
vis2 <- ggplot(ballot.long2, aes(x=days_since_origin, y=rep)) + geom_point(color = "red", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Republican Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
vis2
```

### Reproduced the visualizations as PDF

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggsave(filename = "results/vis1.pdf", plot = vis1, device = "pdf", 
       width = 6, height = 5, units = "in")

ggsave(filename = "results/vis2.pdf", plot = vis2, device = "pdf", 
       width = 6, height = 5, units = "in")


```


### References: 