---
title: "Final Project - SK4.1"
author: "Sarah Im" 
output: html_document
date: "2024-12-04"
bibliography: doc/myrefs.bib
---

# Do voters want Republicans or Democrats in Congress? 

## Background

Public opinion polls from different pollsters and organizations are used to demonstrate the performance and reputation of the political parties in our democratic society. 
The public approval ratings of the Democratic and Republican parties were measured using public opinion polls in this study. The voter behavior in the United States has evolved over the years and can be studied using public opinion polling. These statistics are often used for campaign strategies of different groups and as a tool in academics and journalism (Hillygus, 2011). This study utilizes data from FiveThirtyEight’s generic congressional ballots and focuses on the approval ratings of the democratic and republican parties in congress from November 21, 2020 to August 20, 2022 (FiveThirtyEight, 2022). Analysis of these polls can reveal insights such as the changing voter preference and attitude in correspondence to key events such as election cycles, policy changes, and national crises. However, the credibility and biases over public opinion polling often changes when an individual views the polls. Some individuals view certain polls as “more reliable” if the polls align with their personal beliefs. In addition, correlation between party polarization and poll results can be found (Maddison, Hillygus, 2020).  The interpretation of the poll data can be essential for assessing the political dynamics in the United States and can reveal trends within voter alignment, policy approval, etc. 

This research focuses on the question: 

How have approval ratings between the Democratic and Republican parties in Congress changed over time? 



### Load required libraries:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loading required libraries
library(tidyr)
library(janitor)
library(ggplot2)
library(dplyr)
```


### Importing and Cleaning Raw Data for the analysis of the Democratic Party 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.data <- read.csv("data/generic_polllist (1).csv")

ballot.Approve <- ballot.data %>%
  clean_names() %>%
  select(startdate, dem) %>%
  na.omit()

ballot.Approve$startdate <- as.Date(ballot.Approve$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve$days_since_origin <- as.numeric(ballot.Approve$startdate - origin_date)

ballot.Approve
```
### Further converted data set format for easier plotting 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve
ballot.wide <- pivot_wider(ballot.Approve, names_from = startdate, values_from = dem, values_fn=mean) 
ballot.wide

ballot.long <- pivot_longer(ballot.wide, cols = !days_since_origin, names_to = "startdate", values_to = "dem")
```


### Data was plotted using ggplot to represent the Democratic Approval Rates

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 1. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in approval ratings of the democratic party in congress over time since the origin of the polls on 11/21/20. The democratic party experienced a mean approval rating of around 43.9% with the range being approximately 11.3% (min. value = 38.3%, highest value=50.0%) from 11/21/20 to 11/05/2022."}

vis1 <- ggplot(ballot.long, aes(x=days_since_origin, y=dem)) + geom_point(color = "blue", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Democratic Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + ylim(35,50)
vis1
```

### Cleaning Raw Data for the analysis of the Republican Party 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve2 <- ballot.data %>%
  clean_names() %>%
  select(startdate, rep) %>%
  na.omit()

ballot.Approve2$startdate <- as.Date(ballot.Approve2$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve2$days_since_origin <- as.numeric(ballot.Approve2$startdate - origin_date)
ballot.Approve2
```


### Further converted data set format for easier plotting 

```{r echo=FALSE, message=FALSE, warning=FALSE}
ballot.Approve2
ballot.wide2 <- pivot_wider(ballot.Approve2, names_from = startdate, values_from = rep, values_fn=mean) 
ballot.long2 <- pivot_longer(ballot.wide2, cols = !days_since_origin, names_to = "startdate", values_to = "rep") 
```

### Data was plotted using ggplot to represent the Republican Party Approval Rates

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 2. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in average approval ratings of the republican party in congress over time since the origin of the polls on 11/21/20. The republican party experienced a mean approval rating of around 42.28% with the range being 14.7% (lowest rating = 35.0%, highest value = 49.7%) from 11/21/20 to 11/05/22."}
vis2 <- ggplot(ballot.long2, aes(x=days_since_origin, y=rep)) + geom_point(color = "red", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Republican Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
vis2
```

### Reproduced the visualizations as PDF

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggsave(filename = "results/vis1.pdf", plot = vis1, device = "pdf", 
       width = 6, height = 5, units = "in")

ggsave(filename = "results/vis2.pdf", plot = vis2, device = "pdf", 
       width = 6, height = 5, units = "in")


```
