---
title: "Final Project - SK4.1"
author: "Sarah Im" 
output: html_document
date: "2024-12-04"
bibliography: doc/myrefs.bib
---

# Do voters want Republicans or Democrats in Congress? 

## Background

Public opinion polls from different pollsters and organizations are used to demonstrate the performance and reputation of the political parties in our democratic society. 
The public approval ratings of the Democratic and Republican parties were measured using public opinion polls in this study. The voter behavior in the United States has evolved over the years and can be studied using public opinion polling. These statistics are often used for campaign strategies of different groups and as a tool in academics and journalism [@hillygus2011evolution]. This study utilizes data from FiveThirtyEight’s generic congressional ballots and focuses on the approval ratings of the democratic and republican parties in congress from November 21, 2020 to August 20, 2022 [@fivethirtyeight2022]. Analysis of these polls can reveal insights such as the changing voter preference and attitude in correspondence to key events such as election cycles, policy changes, and national crises. However, the credibility and biases over public opinion polling often changes when an individual views the polls. Some individuals view certain polls as “more reliable” if the polls align with their personal beliefs. In addition, correlation between party polarization and poll results can be found [@madson2020all]. The interpretation of the poll data can be essential for assessing the political dynamics in the United States and can reveal trends within voter alignment, policy approval, etc. 

This research focuses on the question: 

How have approval ratings between the Democratic and Republican parties in Congress changed over time? 

## Analysis 

In this study, the dataset 'generic_polllist(1).csv' was imported from FiveThirtyEight’s generic congressional ballots which contained voter ratings from November 2020 to November 2022. First off, the required libary packsges tidyr, janitor, ggplot2, and dplyr were loaded. The data set was uploaded onto R studio using read.csv, processed, and cleaned by standardizing the column names using clean_names(). Data points for the specific political party (dem or rep) were selected using select(). Missing values were removed from the original data set using na.omit(). New data sets were created per political party for the cleaned data (ballot.Approve and ballot.Approve2) Furthermore, the dates from the original dataset were transformed and formatted to calculate the number of days since the origin date of the polls (11/21/2020) using as.Date and origin_date to allow for better visualization of data trends. For plotting the visualizations, the data was transformed using pivot_wider and pivot_longer to plot the approval ratings of both parties separately. The ggplot2 package was used to plot the data and days since origin versus the approval ratings for each respected political party. The theme and points were adjusted using theme_minimal() and geom_point to promote visibility. The y and x-axis limits were further adjusted for better visualizations. For the visualization for the democratic party, the data points are represented in blue and demonstrate the daily average approval ratings. For the visualization for the republican party, the data points are represented in red and demonstrate the daily average approval ratings. Lastly, ggsave was used to convert the plotted visualizations into pdf format and was placed in results. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loading required libraries
library(tidyr)
library(janitor)
library(ggplot2)
library(dplyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Data Import and Data Cleaning of the Democratic Party 
ballot.data <- read.csv("data/generic_polllist (1).csv")

ballot.Approve <- ballot.data %>%
  clean_names() %>%
  select(startdate, dem) %>%
  na.omit()

ballot.Approve$startdate <- as.Date(ballot.Approve$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve$days_since_origin <- as.numeric(ballot.Approve$startdate - origin_date)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Data Transformation: Conversion of the Data set 
ballot.wide <- pivot_wider(ballot.Approve, names_from = startdate, values_from = dem, values_fn=mean) 

ballot.long <- pivot_longer(ballot.wide, cols = !days_since_origin, names_to = "startdate", values_to = "dem")
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
# Visualization of the data using ggplot to represent the Democratic Approval Rates
vis1 <- ggplot(ballot.long, aes(x=days_since_origin, y=dem)) + geom_point(color = "blue", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Democratic Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + ylim(35,50)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Data Cleaning of the Republican Party 
ballot.Approve2 <- ballot.data %>%
  clean_names() %>%
  select(startdate, rep) %>%
  na.omit()

ballot.Approve2$startdate <- as.Date(ballot.Approve2$startdate, format = "%m/%d/%Y")
origin_date <- as.Date("11/21/2020", format = "%m/%d/%Y")
ballot.Approve2$days_since_origin <- as.numeric(ballot.Approve2$startdate - origin_date)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Data Transformation: Conversion of the Data set  
ballot.wide2 <- pivot_wider(ballot.Approve2, names_from = startdate, values_from = rep, values_fn=mean) 
ballot.long2 <- pivot_longer(ballot.wide2, cols = !days_since_origin, names_to = "startdate", values_to = "rep") 
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Visualization of the data using ggplot to represent the Republican Approval Rates
vis2 <- ggplot(ballot.long2, aes(x=days_since_origin, y=rep)) + geom_point(color = "red", pch=19, alpha=0.5) + labs(x = "Days Since Origin of Polls (11/21/20)", y = "Approval Rating from Polls (%)", title = "Approval of Republican Party in Congress Over Time") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Reproduced the visualizations as PDF
ggsave(filename = "results/vis1.pdf", plot = vis1, device = "pdf", 
       width = 6, height = 5, units = "in")

ggsave(filename = "results/vis2.pdf", plot = vis2, device = "pdf", 
       width = 6, height = 5, units = "in")


```

## Results 

### The Democratic Party 

```{r echo=FALSE, message=FALSE, warning=FALSE,, fig.cap = "Figure 1. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in approval ratings of the democratic party in congress over time since the origin of the polls on 11/21/20. The democratic party experienced a mean approval rating of around 43.9% with the range being approximately 11.3% (min. value = 38.3%, highest value=50.0%) from 11/21/20 to 11/05/2022."}

vis1
```

### The Republican Party

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 2. Plot of Approval Rating from Polls vs. Days Since Origin of polls. Plot demonstrates the change in average approval ratings of the republican party in congress over time since the origin of the polls on 11/21/20. The republican party experienced a mean approval rating of around 42.28% with the range being 14.7% (lowest rating = 35.0%, highest value = 49.7%) from 11/21/20 to 11/05/22."}

vis2
```


## Discussion 

Figure 1 depicts the Democratic Party's approval ratings over time. Based on the data trends and visualizations, the democratic party’s approval was observed to have fluctuated within an approximate range of 12 percent with the highest approval rating of 50% and the lowest approval rating of 38.3%. The average approval rating of the democratic party was approximately 43.9%. Figure 2 depicts the Republican Party's approval ratings overtime. The republican party’s approval ratings exhibited a broader range with the highest approval rating of 49.7% and the lowest approval rating of 35.0%. The average approval rating of the republican party was approximately 42.28%. 

## Conclusion 

Overall, both political parties exhibited similar average approval ratings with the democratic party demonstrating slightly higher average approval ratings over time. The democratic party sustained a higher average approval rating (43.9%) compared to the republican party (42.28%). However, there are not enough substantial differences between the two parties that establishes a clear preference of one party over the other during the time period. 


### References: 